Developmental trajectories of word recognition: a priming study in bilinguals

### 2023-07-17 manuscript/manuscript.qmd
Update docs

---
title: "Developmental trajectories of word recognition: a priming study in bilinguals"
editor: source
author:
  - name: Gonzalo Garcia-Castro
    orcid: 0000-0002-8553-4209
    email: gonzalo.garciadecastro@upf.edu
    affiliations:
      - ref: upf
    corresponding: true
  - name: Serene Siow
    affiliations:
      - ref: oxf
  - name: Núria Sebastian-Galles
    affiliations:
      - ref: upf
  - name: Kim Plunkett
    affiliations:
      - ref: oxf
affiliations:
  - id: upf
    name: Center for Brain and Cognition, Universitat Pompeu Fabra
  - id: oxf
    name: Department of Experimental Psychology, University of Oxford

pagetitle: Cross-linguistic similarity and  word acquisition
abstract: "Previous studies have provided evidence that lexical access is language non-selective in bilinguals: recognising and producing words in one language activates lexical representations of words in the other language (e.g., Costa et al., 2000; Thierry & Wu, 2007). It has been suggested that this parallel activation is already present during toddlerhood (Jardak & Byers-Heinlein, 2019; Von Holzen & Mani, 2012), but it is unclear how it impacts the developing lexicon. In the present study, we tested Catalan-Spanish and English-Spanish bilinguals in a word recognition task, in which they were presented with words in their dominant language exclusively. In each trial, two pictures were presented side-by-side and one of them was named (target picture). Participants' visual preference for the named picture was taken as an indicator of word recognition. Each pair of pictures was preceded by the presentation of a silent image (prime picture). We designed three types of trials: (1) cognate trials in which prime and target labels shared phonological onset and the prime label was a cognate (e.g., flower-flor / fork), (2) non-cognate trials, in which prime and target labels shared phonological onset but the prime label was a non-cognate (e.g., frog-rana / fork), and (3) unrelated trials, in which prime and target labels did not share onset (unrelated trials; car-coche / fork). In line with previous studies, we expected participants to generate implicit labels for the prime pictures, which should interfere with target word recognition when both words share phonological onset (e.g., Mani & Plunkett, 2011). Additionally, under the hypothesis that bilingual participants would activate labels for the prime pictures in both languages, we predicted that interference would be stronger after cognate primes (labels from both languages share phonological overlap with target word recognition) than after non-cognate primes (only the label in the target language share phonological overlap with target word recognition). We tested bilinguals (n = 46) and same-aged monolingual controls, n = 123) at three age points (21, 25, and 30 months) to investigate how any cross-language priming effect emerged or changed across these ages."
thanks: "This research was supported by grants from the Spanish Ministerio de Ciencia, Innovación y Universidades (PGC2018-101831-B-I00 and PRE2019-088165), and the Catalan Government [ICREA (Catalan Institution for Research and Advanced Studies) Academia 2019 award]. Gonzalo Garcia-Castro was supported by a fellowship of the Spanish Ministerio de Ciencia, Innovación y Universidades (FPI 2019). The authors declare no conflicts of interest with regard to the funding source of this study. We thank Xavier Mayoral, Silvia Blanch, Cristina Cuadrado, and Irina Lepadatu for their technical support. We also thank Cristina Dominguez, Katia Pistrin, and Nicola Gillen for their efforts in recruiting infants. We would like to thank the clinics Quirón and Sagrada Familia that allowed us to recruit participants in their premises. We also thank all families and infants who participated in the experiments. Data collection was half-way when the COVID-19 pandemic started. We would like to pay special tribute to the families that collaborated with us under these difficult circumstances."
keywords: word recognition, cognate, vocabulary, bilingualism, eye-tracking, growth curve analysis, bayesian
csl: assets/bibliography/apa7.csl
bibliography: assets/bibliography/references.bib
format:
  pdf:
    pdf-engine: lualatex
    number-sections: true
    whitespace: small
  docx:
  	reference-doc: assets/templates/template.docx
linestretch: 1.5
toc: false
fig-dpi: 1000
echo: false
warning: false
---

```{r}
#| label: setup
#| include: false

tar_load_everything() # load target objects

# load packages
library(dplyr)
library(tidyr)
library(stringr)
library(tibble)
library(purrr)
library(lubridate)
library(ggplot2)
library(gt)
library(tidybayes)
library(brms)
library(polypoly)
library(tidytext)
library(janitor)
library(forcats)
library(scales)
library(patchwork)
library(bvq)

# set ggplot theme and colour palette
theme_set(theme_custom()) # set custom ggplot theme
clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")

options(ggplot2.ordinal.fill = clrs[c(1, 4, 5)],
		ggplot2.ordinal.colour = clrs[c(1, 4, 5)],
		ggplot2.discrete.fill = clrs[c(1, 4, 5)],
		ggplot2.discrete.colour = clrs[c(1, 4, 5)],
		ggplot2.continuous.fill = ggplot2::scale_color_gradient,
		ggplot2.continuous.colour = ggplot2::scale_color_gradient)

# prepare data
attrition <- attrition %>% 
	mutate(
		valid_gaze = valid_gaze_prime & valid_gaze_target,
		valid_status = case_when(
			!valid_gaze ~ "Invalid gaze",
			!valid_participant ~ "Invalid participant",
			TRUE ~ "Valid"
		))

participants <- participants %>% 
	left_join(distinct(attrition, id, age_group, valid_participant)) %>% 
	drop_na(id) %>% 
	left_join(vocabulary) %>% 
	mutate(vocab_size_l1_center = scale(vocab_size_l1)[,1])

@@

All materials, data, and reproducible code can be found at the OSF ([https://osf.io/hy984/](https://osf.io/ckydb/)) and GitHub ([https://github.com/gongcastro/cognate-priming](https://github.com/gongcastro/cognate-priming)) repositories. This study was conducted according to guidelines laid down in the Declaration of Helsinki, and was approved by the Drug Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar, reference 2020/9080/I.

@@

```{r}
#| label: participants-numbers
n_participants_total <- length(unique(participants$id))
n_participants_sessions <- count(participants, id, name = "n_sessions") %>% 
	count(n_sessions) %>% 
	split(.$n_sessions)
n_sessions_total <- count(participants)
n_sessions_age_group <- participants %>% 
	group_by(age_group) %>% 
	summarise(across(age, lst(mean, sd, min, max)),
			  n = n(),
			  .groups = "drop") %>% 
	mutate(across(age_mean:age_max, round, 2)) %>% 
	split(f = .$age_group) %>% 
	set_names(c("age_21", "age_25", "age_30"))
n_sessions_dominance <- count(participants, test_language) %>% 
	split(f = .$test_language) %>% 
	set_names(c("catalan", "spanish"))
n_sessions_dominance_age_group <- count(participants, age_group, test_language) %>% 
	split(f = .$test_language) %>% 
	set_names(c("catalan", "spanish")) %>% 
	map(~split(., f = .$age_group) %>% 
			set_names(c("age_21", "age_25", "age_30")))
n_sessions_lp <- participants %>% 
	count(lp) %>% 
	split(f = .$lp) %>% 
	set_names(c("monolingual", "bilingual"))
n_sessions_lp_age_group <- participants %>% 
	count(lp, age_group) %>% 
	split(f = .$lp) %>% 
	set_names(c("monolingual", "bilingual")) %>% 
	map(~split(., f = .$age_group) %>% 
			set_names(c("age_21", "age_25", "age_30")))
```

@@

We assessed participants' language profile using the Language Exposure Questionnaire [LEQ, @bosch2001evidence]. Before each experimental session, the experimenter asked the caretakers to estimate the amount of hours per day they and other people in the infant's social circle have spent speaking to the infant in any language since birth. The output of this interview is an estimated degree of exposure (DoE) to each language, indicated by the proportion of time the infant was reported to have listened to each language. According to this estimate, we classified participants as Catalan- or Spanish-dominant if the language with highest DoE was Catalan or Spanish, respectively, and tested the participant in the stimuli set that contained words in their native language. We collected data from `r n_sessions_dominance$catalan$n` Catalan-dominant participants in Catalan (`r n_sessions_dominance_age_group$catalan$age_21$n` at 21 months, `r n_sessions_dominance_age_group$catalan$age_25$n` at 25 months, and `r n_sessions_dominance_age_group$catalan$age_30$n` at 30 months). We further classified participants as monolinguals if the DoE to their dominant language exceeded 80% of the total DoE to Catalan and Spanish, and as bilinguals otherwise. Participants with DoE to language other than Catalan or Spanish were excluded from analyses. This divided the sample into `r n_sessions_lp$monolingual$n` monolinguals (`r n_sessions_lp_age_group$monolinguals$age_21` at 21 months, `r n_sessions_lp_age_group$monolinguals$age_25` at 25 months, and `r n_sessions_lp_age_group$monolinguals$age_30` at 30 months), and `r n_sessions_lp$bilingual$n` bilinguals (`r n_sessions_lp_age_group$monolinguals$age_21` at 21 months, `r n_sessions_lp_age_group$bilingual$age_25` at 25 months, and `r n_sessions_lp_age_group$bilingual$age_30` at 30 months). tbl-participants-lp summarises the linguistic profile of our sample.

@@

>>>>

---

---

```{r setup}
#| label: setup
#| echo: false
#| message: false
#| warning: false
# load objects -----------------------------------------------------------------

targets::tar_config_set(store = here::here('_targets'),
						script = here::here('_targets.R'))

targets::tar_load_globals()

targets::tar_load(c(participants, stimuli, vocabulary,
					attrition_trials, attrition_participants,
					gaze_aoi,
					data_time, data_summary))

library(knitr)
library(kableExtra)
library(ggplot2)
library(gt)

# set ggplot theme and colour palette ------------------------------------------

my_theme <- theme_minimal() +
	theme(panel.grid = element_line(colour = "grey",
									linetype = "dotted"),
		  axis.line = element_line(colour = "black"),
		  text = element_text(size = 12, colour = "black"),
		  axis.text = element_text(colour = "black"))

theme_set(my_theme)

set.seed(888)

# prepare data
attrition_participant <- attrition_participants |> 
	mutate(
		valid_status = case_when(
			!is_valid_gaze ~ "Invalid gaze",
			!is_valid_participant ~ "Invalid participant",
			TRUE ~ "Valid"
		))

participants <- participants |> 
	left_join(distinct(attrition_participants,
					   id, age_group, is_valid_participant)) |> 
	drop_na(id) |> 
	left_join(vocabulary) |> 
	mutate(prop_l1_c = scale(prop_l1)[,1])

@@

All materials, data, and reproducible code can be found at the OSF ([https://osf.io/hy984/](https://osf.io/ckydb/)) and GitHub ([https://github.com/gongcastro/cognate-priming](https://github.com/gongcastro/cognate-priming)) repositories. This study was conducted according to guidelines laid down in the Declaration of Helsinki, and was approved by the Drug Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar, reference 2020/9080/I. Before every testing session, caretakers were asked to read and sign an informed consent form, and were given a small gift at the end of it

@@

```{r}
#| label: participants-numbers
n_participants_total <- length(unique(participants$id))
n_participants_sessions <- count(participants, id, name = "n_sessions") |> 
	count(n_sessions) |> 
	split(.$n_sessions)
n_sessions_total <- count(participants)
n_sessions_age_group <- participants |> 
	group_by(age_group) |> 
	summarise(across(age, lst(mean, sd, min, max)),
			  n = n(),
			  .groups = "drop") |> 
	mutate(across(age_mean:age_max, round, 2)) |> 
	split(f = .$age_group) |> 
	set_names(c("age_21", "age_25", "age_30"))
n_sessions_dominance <- count(participants, test_language) |> 
	split(f = .$test_language) |> 
	set_names(c("catalan", "spanish"))
n_sessions_dominance_age_group <- count(participants, age_group, test_language) |> 
	split(f = .$test_language) |> 
	set_names(c("catalan", "spanish")) |> 
	map(~split(., f = .$age_group) |> 
			set_names(c("age_21", "age_25", "age_30")))
n_sessions_lp <- participants |> 
	count(lp) |> 
	split(f = .$lp) |> 
	set_names(c("monolingual", "bilingual"))
n_sessions_lp_age_group <- participants |> 
	count(lp, age_group) |> 
	split(f = .$lp) |> 
	set_names(c("monolingual", "bilingual")) |> 
	map(~split(., f = .$age_group) |> 
			set_names(c("age_21", "age_25", "age_30")))
```

@@

We assessed participants' language profile using the Language Exposure Questionnaire [LEQ, @bosch2001evidence]. Before each experimental session, the experimenter asked the caretakers to estimate the amount of hours per day they and other people in the infant's social circle have spent speaking to the infant in any language since birth. The output of this interview is an estimated degree of exposure (DoE) to each language, indicated by the proportion of time the infant was reported to have listened to each language. According to this estimate, we classified participants as Catalan- or Spanish-dominant if the language with highest DoE was Catalan or Spanish, respectively, and tested the participant in the stimuli set that contained words in their native language. We collected data from `r n_sessions_dominance$catalan$n` Catalan-dominant participants in Catalan (`r n_sessions_dominance_age_group$catalan$age_21$n` at 21 months, `r n_sessions_dominance_age_group$catalan$age_25$n` at 25 months, and `r n_sessions_dominance_age_group$catalan$age_30$n` at 30 months). We further classified participants as monolinguals if the DoE to their dominant language exceeded 80% of the total DoE to Catalan and Spanish, and as bilinguals otherwise. Participants with DoE to language other than Catalan or Spanish were excluded from analyses. This divided the sample into `r n_sessions_lp$monolingual$n` monolinguals (`r n_sessions_lp_age_group$monolinguals$age_21` at 21 months, `r n_sessions_lp_age_group$monolinguals$age_25` at 25 months, and `r n_sessions_lp_age_group$monolinguals$age_30` at 30 months), and `r n_sessions_lp$bilingual$n` bilinguals (`r n_sessions_lp_age_group$monolinguals$age_21` at 21 months, `r n_sessions_lp_age_group$bilingual$age_25` at 25 months, and `r n_sessions_lp_age_group$bilingual$age_30` at 30 months). @tbl-participants-lp shows a detailed description of the linguistic profile of our sample.

```{r}
#| label: tbl-participants-lp
#| tbl-cap: "Description of language profile of test participants. Data are summarised for each age group, and for monolinguals and bilinguals separately."
participants |> 
	left_join(attrition_participants,
			  by = join_by(id, 
			  			 age_group,
			  			 is_valid_participant)) |> 
	filter(is_valid_participant) |> 
	select(id, age_group, age, lp,
		   doe_catalan, doe_spanish, test_language) |> 
	mutate(id = paste0(id, " (", age_group, ")")) |>
	add_count(lp, 
			  name = "n_lp") |> 
	add_count(age_group, 
			  name = "n_age_group") |> 
	pivot_longer(starts_with("doe_"),
				 names_to = "language",
				 values_to = "doe") |>
	add_count(age_group,
			  test_language,
			  name = "n_age_test") |> 
	mutate(language = str_to_sentence(str_remove_all(language, "doe_")),
		   age_group = paste0(age_group, " (N = ", n_age_group, ")"),
		   test_language = paste0("Tested in ", test_language, 
		   					   " (N = ", n_age_test, ")"),
		   lp = factor(lp, levels = rev(unique(lp))))  |> 
	summarise(across(c(doe, age), lst(mean, sd)),
			  .by = c(age_group, lp, test_language, language)) |> 
	pivot_wider(id_cols = c(age_group, test_language),
				names_from = c(language, lp),
				values_from = c(matches("doe"), age_mean, age_sd),
				names_repair = janitor::make_clean_names) |> 
	relocate(age_group, test_language,
			 matches("monolingual"),
			 matches("bilingual")) |> 
	select(-c(age_mean_spanish_monolingual,
			  age_mean_spanish_bilingual,
			  age_sd_spanish_monolingual,
			  age_sd_spanish_bilingual)) |> 
	arrange(age_group, test_language) |> 
	gt(rowname_col = "test_language", 
	   groupname_col = "age_group", 
	   row_group.sep = ": ") |> 
	tab_spanner(md("Monolingual (*N* = 162)"), matches("monolingual")) |> 
	tab_spanner(md("Bilingual (*N* = 133)"), matches("bilingual")) |>
	fmt_number(matches("doe"), decimals = 1, scale_by = 100) |> 
	fmt_number(matches("age"), decimals = 1) |> 
	cols_merge_uncert(col_val = age_mean_catalan_monolingual, 
					  col_uncert = age_sd_catalan_monolingual) |> 
	cols_merge_uncert(col_val = age_mean_catalan_bilingual, 
					  col_uncert = age_sd_catalan_bilingual) |> 
	cols_merge_uncert(col_val = doe_mean_catalan_monolingual,
					  col_uncert = doe_sd_catalan_monolingual) |> 
	cols_merge_uncert(col_val = doe_mean_catalan_bilingual,
					  col_uncert = doe_sd_catalan_bilingual) |> 
	cols_merge_uncert(col_val = doe_mean_spanish_monolingual, 
					  col_uncert = doe_sd_spanish_monolingual) |> 
	cols_merge_uncert(col_val = doe_mean_spanish_bilingual, 
					  col_uncert = doe_sd_spanish_bilingual) |> 
	cols_label(age_mean_catalan_monolingual = "Age (months)",
			   age_mean_catalan_bilingual = "Age (months)",
			   doe_mean_catalan_monolingual = "Catalan (%)",
			   doe_mean_catalan_bilingual = "Catalan (%)",
			   doe_mean_spanish_monolingual = "Spanish (%)",
			   doe_mean_spanish_bilingual = "Spanish (%)") |> 
	tab_style(cell_text(weight = "bold"),
			  list(cells_column_spanners())) |> 
	tab_style(cell_text(size = "medium"),
			  list(cells_body(),
			  	 cells_stub()))
```

We collected vocabulary data using parental responses to the Barcelona Vocabulary Inventory [BVQ, @garcia-castro2023bvq], an online vocabulary checklist inspired in several adaptations of the the Communicative Developmental Inventory [CDI, @fenson2004variability] developed to assess the vocabulary size of Catalan-Spanish bilingual toddlers. Families received a link to the BVQ immediately after each experimental session, and were given two weeks to fill it. Two weeks after being sent the link, `r table(vocabulary$is_imputed)[2]` (`r scales::percent(table(vocabulary$is_imputed)[2]/nrow(vocabulary))` families failed to provide a complete response to the BVQ. We imputed their vocabulary size scores using multiple imputation [@vanbuuren2018flexible]. We used predictive mean matching to impute missing scores using participants' information about their age group and language profile. We obtained several receptive and expressive vocabulary size measures from each participant's vocabulary: Catalan vocabulary size (proportion of the words in the Catalan checklist reported as acquired), Spanish vocabulary size (proportion of the words in the Catalan checklist reported as acquired), and total vocabulary size (proportion of the words in both checklists reported as acquired). @tbl-participants-vocabulary shows a detailed description of participants' observed and imputed vocabulary size scores.

```{r tbl-participants-vocabulary}
participants |> 
	mutate(is_imputed = ifelse(is_imputed, "Imputed", "Observed"),
		   cat_prop = ifelse(test_language=="Catalan",
		   				  l1_prop, 
		   				  l2_prop),
		   spa_prop = ifelse(test_language=="Spanish",
		   				  l1_prop,
		   				  l2_prop)) |> 
	select(id, age_group, lp, matches("prop"), is_imputed,
		   -c(matches("l1|l2"), starts_with("te_"))) |> 
	summarise(across(matches("prop"), lst(mean, sd)),
			  .by = c(age_group, lp, is_imputed)) |> 
	pivot_wider(names_from = is_imputed,
				values_from = matches("prop"),
				names_repair = janitor::make_clean_names) |> 
	select(-matches("concept")) |> 
	arrange(age_group, lp) |> 
	gt(rowname_col = "lp",
	   groupname_col = "age_group") |> 
	fmt_number(matches("prop"), 
			   decimals = 1,
			   scale_by = 100) |> 
	cols_merge(c(total_prop_mean_observed,
				 total_prop_sd_observed), 
				 pattern = "{1} ({2})") |> 
	cols_merge(c(total_prop_mean_imputed, 
				 total_prop_sd_imputed),
				 pattern = "{1} ({2})") |> 
	cols_merge(c(cat_prop_mean_observed,
				 cat_prop_sd_observed),
				 pattern = "{1} ({2})") |> 
	cols_merge(c(cat_prop_mean_imputed, 
				 cat_prop_sd_imputed),
				 pattern = "{1} ({2})") |> 
	cols_merge(c(spa_prop_mean_observed,
				 spa_prop_sd_observed),
				 pattern = "{1} ({2})") |> 
	cols_merge(c(spa_prop_mean_imputed,
				 spa_prop_sd_imputed),
				 pattern = "{1} ({2})") |> 
	tab_spanner("Total", matches("total")) |> 
	tab_spanner("Catalan", matches("cat")) |> 
	tab_spanner("Spanish", matches("spa")) |> 
	cols_label(total_prop_mean_observed = "Observed",
			   cat_prop_mean_observed = "Observed",
			   spa_prop_mean_observed = "Observed",
			   total_prop_mean_imputed = "Imputed",
			   cat_prop_mean_imputed = "Imputed",
			   spa_prop_mean_imputed = "Imputed") |> 
	tab_style(cell_text(weight = "bold"),
			  list(cells_column_spanners())) |> 
	tab_style(cell_text(size = "medium"),
			  list(cells_body(),
			  	 cells_stub()))
```

@@

### Data processing

```{r attrition-numbers}
n_trials <- nrow(attrition_trials)
n_sessions <- distinct(attrition_trials, id, age_group) |> 
	nrow()
n_participants <- distinct(attrition_trials, id) |> 
	nrow()

n_trials_prime <- attrition_trials |> 
	filter(is_valid_gaze_prime)

n_trials_test <- n_trials_prime |> 
	filter(is_valid_gaze_test)

n_trials_each <- n_trials_test |> 
	filter(is_valid_gaze_test_each)

n_trials_participant <- n_trials_each |> 
	inner_join(filter(attrition_participants,
					  is_valid_participant),
			   by = join_by(id, age_group))

n_sessions_valid <- attrition_participants |> 
	filter(is_valid_participant)

n_participants_valid <- n_trials_participant |> 
	distinct(id)

n_longitudinal <- count(n_sessions_valid, id, name = "sessions") |> 
	count(sessions)

n_participants_include <- read_csv("data-raw/participants.csv",
								   show_col_types = FALSE)
```

We gathered data for `r format(n_trials, big.mark = ",")` trials from `r n_sessions` experimental sessions, provided by `r n_participants` distinct participants. Missing eye-tracker samples were interpolated using the last-observation-carried-forward [see @zettersten2022peekbank for a similar approach]. We excluded trials in which the participant did not fixate the prime at least half of the time of the prime phase ($\geq$750 ms, *n* = `r format(n_trials - nrow(n_trials_prime), big.mark = ",")`), or the target and distractor during the test phase ($\geq$1,000 ms, *n* = `r format(nrow(n_trials_prime) - nrow(n_trials_test), big.mark = ",")`). Additionally, we also excluded trials in which the participant failed to fixate both target *and* distractor pictures for at least 100 ms (5% of the time, `r format(nrow(n_trials_test) - nrow(n_trials_each), big.mark = ",")`). Finally we excluded trials from experimental sessions in which the participant did not provide at least two valid trials in each of the three experimental conditions (cognate prime, non-cognate prime, and unrelated prime) which resulted in the exclusion of `r nrow(distinct(n_trials_each, id, age_group)) - nrow(n_sessions_valid)` experimental sessions and their `r nrow(n_trials_each) - nrow(n_trials_participant)` remaining trials. The final data set comprised `r format(sum(data_summary$.ntrials), big.mark = ",")` from `r n_sessions_valid` experimental sessions, provided by `r n_participants_valid` distinct participants. Of those participants, `r n_longitudinal[1]` provided data from one experimental session, `r n_longitudinal[2]` provided data from two experimental sessions, and `r n_longitudinal[1]` provided data from three experimental sessions.

We defined our time window of interest from 300 ms after the onset of the test phase (target and distractor presentation) until the end of the test phase (2,000 ms). For each trial, we chunked the time domain into 17 time bins of 100 ms of duration. We then calculated, for each experimental session, time bin, and condition, participant's proportion of target and distractor fixations. Finally, we computed the empirical log-odds of target and distractor fixations, which we introduced in the statistical analyses as our response variable.

```{r} 
#| label: tbl-attrition-group
#| tbl-cap: "Summary of the dataset."
attrition_trials |> 
	left_join(select(participants, age_group, lp, id, date_test),
			  by = join_by(id, age_group)) |> 
	left_join(attrition_participants,
			  by = join_by(id, age_group)) |> 
	add_count(age_group, lp, name = "n_trials") |> 
	summarise(valid_trial_count = sum(is_valid_trial, na.rm = TRUE),
			  .by = c(id, lp, trial_type, date_test,
			  		age_group, is_valid_participant, n_trials)) |> 
	pivot_wider(names_from = trial_type, 
				values_from = valid_trial_count, 
				values_fill = 0,
				names_repair = janitor::make_clean_names) |> 
	rename(noncognate = non_cognate) |> 
	add_count(lp, age_group) |>
	summarise(across(c(cognate, noncognate, unrelated),
					 lst(sum, mean, sd)),
			  n_valid = sum(is_valid_participant),
			  .by = c(lp, age_group, n, n_trials)) |>
	relocate(n_valid, .after = n) |> 
	mutate(n_prop = n_valid/n) |> 
	arrange(age_group, lp) |> 
	relocate(age_group, lp, n_valid, n_prop, 
			 matches("cognate"),
			 matches("noncognate"),
			 matches("unrelated")) |> 
	gt(groupname_col = "age_group",
	   rowname_col = "lp")  |>
	cols_hide(c(n, n_trials)) |> 
	fmt_number(matches("_mean|_sd"), 
			   decimals = 1) |>
	fmt_percent(matches("_prop"),
				decimals = 1) |> 
	cols_merge_n_pct(col_n = n_valid, 
					 col_pct = n_prop,
					 autohide = TRUE)  |> 
	tab_spanner("Non-cognate", matches("noncognate")) |>
	tab_spanner("Cognate", starts_with("cognate")) |>
	tab_spanner("Unrelated", matches("unrelated")) |>
	tab_spanner("Related", matches("cognate")) |> 
	cols_merge_uncert(col_val = cognate_mean,
					  col_uncert = cognate_sd) |> 
	cols_merge_uncert(col_val = noncognate_mean,
					  col_uncert = noncognate_sd) |> 
	cols_merge_uncert(col_val = unrelated_mean, 
					  col_uncert = unrelated_sd) |> 
	cols_label(age_group = "Age",
			   n_valid = "Participants (included)",
			   cognate_sum = "Trials",
			   cognate_mean = "Mean ± SD",
			   noncognate_sum = "Trials",
			   noncognate_mean = "Mean ± SD",
			   unrelated_sum = "Trials",
			   unrelated_mean = "Mean ± SD") |> 
	summary_rows(groups = TRUE,
				 columns = where(is.integer),
				 fns = list("Sum" = "sum"),
				 fmt = fmt_number,
				 decimals = 0) |>
	summary_rows(groups = TRUE,
				 columns = c(where(is.numeric), -n_valid),
				 fns = list("Mean" = "mean"),
				 fmt = fmt_number,
				 decimals = 1) |>
	grand_summary_rows(columns = where(is.integer),
					   fns = list("N" = "sum"),
					   formatter = fmt_integer) 
```
